Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-02 thesis/acm/paper.tex
Other changes

\category{D.2.7}{Software Engineering}{Distribution, Maintenance, Enhancement}[Restructuring, reverse engineering, and reengineering]
\category{K.6.3}{Computing Milieux}{Software Management}[Software Maintenance]

@@

\section{Introduction} \label{introduction}
%API Design (and a little bit of ecosystems, and a bit of problem statement)
As the software industry and the open-source movement steadily grow, the number of public APIs is increasing. APIs can improve the development speed \cite{stylos2006comparing}, contribute to higher quality software \cite{stylos2006comparing} and increase the reusability of software \cite{afonso2012evaluating}. There is a consensus in that modifications to APIs could negatively impact the users of the API \cite{google_talk} \cite{mcdonnell2013empirical} \cite{robbes2012developers} \cite{henning2007api}. The main reason is that it will require the users to update the code, which is using the API, thus causing a disruption in the application's software ecosystem \cite{messerschmitt2005software}. The most common changes to APIs occur from refactoring \cite{dig2005role} \cite{xing2006refactoring}. 
%Software Evolution 
In the discipline of software evolution, such updates to software are studied from an evolutionary standpoint. Software can be updated for different reasons and these motives can be grouped into corrective, adaptive, perfective and preventive changes \cite{lientz1980software}.

%Gap in Literature
Studies that explore the intent behind API modifications exist \cite{hou2011exploring}. There are, however, no studies that attempt to map the motives behind changes to existing software evolution theories, such as the four categories of maintenance \cite{lientz1980software}. The area of programming language APIs is well-explored \cite{hou2011exploring} \cite{shi2011empirical}, but few studies that explore platform APIs exist \cite{robbes2012developers}. To the best of our knowledge, no studies that explore the motives behind or the effects of API evolution have previously been performed on embedded platform APIs.

@@

\section{Existing Literature} \label{existing_literature}
This section will introduce the topics of API design and software evolution, and provide a review of studies that are directly related to our study.

@@

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field which studies the application of software maintenance activities and processes on software, and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s. Well-known studies from this era established the importance of software maintenance \cite{lientz1980software}. One study would go on to have a particularly profound effect on software maintenance by establishing a set of four categories which describe the intents behind change to software \cite{lientz1980software}. These categories were refined in a further study, in which the number of categories were increased from four to twelve \cite{chapin2001types}. The categories proposed were: Enhancive, corrective, reductive, adaptive, performance, preventive, groomative, updative, reformative, evaluative, consultive and training \cite{chapin2001types}. Each category describes a type of software maintenance activity which relate to one of three particular areas; the code, the software and the customer-experienced functionality \cite{chapin2001types}. This study also helped to further the distinction between software maintenance and software evolution. The study suggested that software maintenance is a part of software evolution and that the two are not interchangeable \cite{chapin2001types}.

@@

\section{Case Company Description} \label{case_company_description}
The case company is a medium-sized company in the domain of security cameras. The company has its headquarters in Lund, Sweden, but it operates worldwide. The company is the global market leader in the markets of network cameras and video encoders. The company develops embedded software for the security cameras, and some of their products are designed to be accessible through APIs that are also developed by the company.

@@

The case was selected using convenience sampling \cite{flyvbjerg2006five}, and a maximum variation strategy was used \cite{benbasat1987case}. As such, the goal was to include as many companies as possible in this study.

@@

\section{Methodology} \label{methodology}
%Intro and summary
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. The goal of the study was twofold. Firstly, the aim was to empirically \textit{explore} the motives behind changes to APIs, and secondly, our ambition was to \textit{validate} the results' level of correctness by comparing our findings with the case company's explanation of their motives. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis.

%Reasons for choosing case study
The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case} \cite{runeson2009guidelines} \cite{yin2013case} \cite{robson2002real}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of motives behind API change, and that input from the industry is vital to the success of the research; especially to be able to answer \textbf{RQ2}.

@@

%Description of code analysis
More specifically, the approach undertaken was divided into three steps where the data was first coded based on which category of maintenance \cite{chapin2001types} the motive of the change corresponded to. The data were then mapped and coded to the laws presented by \cite{lehman1980programs}. The codes were then analysed from a high-level point of view, the main goal of which was to identify trends in the codes. The final step involved analysis of these trends, and it was during this stage the hypotheses were generated and formulated. The same approach taken to reduce risks of research bias and human error in the data collection was performed during results analysis.

@@

\subsection{Continuing Change}
The case company described that including new functional content was one of the top priorities for the 2.0 release. This is supported by the characteristic of the changes identified in this study. 4 of the 13 changes were additions of new modules, and in the re-implemented modules a lot of functional content was added, too. It was also mentioned that the API users had previously requested more frequent updates to the API, despite the drawbacks this could have. One of the main reasons for continuously adapting the API is also to force (!!!wording) the API users to use the API in the way that it is intended. Previously, many applications which worked around the lacking (!!!negative wording) functionality had been developed. These applications were fully functional, but were developed very differently from user to user (!!!wording). We therefore conclude that the law of continuing change holds for APIs.

@@

\subsection{Self Regulation} \label{law3}
When version 1.4 was introduced, no new functionality or re-structuring was introduced. Version 1.4 only added support for additional build platforms. When comparing the changes introduced in 2.0 to the ones previously introduced, it is clear that the growth of the API has increased gradually. During the interview it was also mentioned that the future plans for the API did not include any major changes, and that only a few minor changes were planned in the near future. The future updates to the API will also be made incrementally, and there will not be a big bang update similar to the one of 2.0. This supports that the law of self regulation applies to APIs, assuming that the current stage of the life-cycle of the API can be considered as the centre of the normal distribution curve of the APIs growth.

\subsection{Conservation of Organisational Stability}
The changes to the ACAP API has been made very sporadically, where only two new versions have been released since the deployment of version 1.0. As mentioned in Section \ref{law3} the changes introduced in version 1.4 did not involve any changes to the existing API functions. Consequently, the only significant update to the API was made in the transition to version 2.0. The internal development of the API has also been made sporadically, and there have been periods where the API has not been actively developed. This suggests that the law of conservation of organisational stability does not apply to APIs.

@@

\subsection{Continuing Growth}
Large parts of the changes between the versions were strictly additions of new functional content. 4 new modules were added, and significant additions of new functionality were added to the modules which were re-structured from version 1.4. The interviewee also made it clear that increasing the functionality was one of the major goals with the 2.0 update. This also ties in with the ambition of controlling the way the API users use the API, as mentioned previously. It was also mentioned that the case company had received feedback from the API users that more frequent updates to the functional content was desirable, even though this may require them to update their application code. Because of this, we conclude that the law of continuing growth applies to APIs.

@@

\subsection{Feedback System}
One of the future goals of improvement of the case company is to better include the API users and end-users of the applications in their feedback loop. Currently, the only the API users are included in this feedback loop, and the API architect expressed the desire to also include the end-users in this loop. The reasoning for this was to increase the quality, and to be able to deliver content which is useful for the users, and which leads to good applications being developed. One example of how the API users have been included in the feedback loop, was prior to version 2.0 being released, when API users implemented functionality related to sound, even though there was no interface for this in the API. The experiences of this API user, were later considered when developing the axsound module. One benefit of having a more structured feedback loop, would be that the API developers would be more and faster aware of potential errors in the implementation of the API. Another ambition which was mentioned, was to bring scientific theories and best practices from academia into consideration when developing the API. This would aim to complement the current strategy which is largely based around the expertise and experiences of the API architects. Because of these factors and clearly expressed ambitions, we conclude that the law of feedback system is true for APIs.

@@

RQ1.1 - To what extent can we reverse engineer API change decisions?
Roughly 78\% of our categorisation matched the actual categorisation of the changes. While it is true that we did not manage to successfully categorise every single change, we did categorise a high enough percentage to form a solid conclusion; it is possible to reverse engineer the majority of change decisions using this categorisation technique.

For future studies, this technique could be improved by performing more actions along side the code review. For example, having access to commit logs, design meetings and other such documentation would allow a much more thorough determination of intents to be performed.

RQ2
What this means for industry:
	Greater emphasis should be placed upon communicating with the end users / application developers. APIs should be made with maintainability and extendibility in mind as they seem to be what changes the most between versions.

@@

What this means for academia:

>>>>

\category{D.2.7}{Software Engineering}{Distribution, Maintenance, Enhancement}[Restructuring, reverse engineering and reengineering]
\category{K.6.3}{Computing Milieux}{Software Management}[Software Maintenance]

@@

\section{Introduction} \label{introduction}
%API Design (and a little bit of ecosystems and a bit of problem statement)
As the software industry and the open-source movement steadily grow, the number of public APIs is increasing. APIs can improve the development speed \cite{stylos2006comparing}, contribute to higher quality software \cite{stylos2006comparing} and increase the reusability of software \cite{afonso2012evaluating}. There is a consensus in that modifications to APIs could negatively impact the users of the API \cite{google_talk} \cite{mcdonnell2013empirical} \cite{robbes2012developers} \cite{henning2007api}. The main reason is that it will require the users to update the code, which is using the API, thus causing a disruption in the application's software ecosystem \cite{messerschmitt2005software}. The most common changes to APIs occur from refactoring \cite{dig2005role} \cite{xing2006refactoring}. 
%Software Evolution 
In the discipline of software evolution, such updates to software are studied from an evolutionary standpoint. Software can be updated for different reasons and these motives can be grouped into corrective, adaptive, perfective and preventive changes \cite{lientz1980software}.

%Gap in Literature
Studies that explore the intent behind API modifications exist \cite{hou2011exploring}. There are, however, no studies that attempt to map the motives behind changes to existing software evolution theories, such as the four categories of maintenance \cite{lientz1980software}. The area of programming language APIs is well-explored \cite{hou2011exploring} \cite{shi2011empirical}, but few studies that explore platform APIs exist \cite{robbes2012developers}. To the best of our knowledge, no studies that explore the motives behind or the effects of API evolution have previously been performed on embedded platform APIs. Reverse engineering design decisions is a challenging yet interesting problem. The better we can understand why changes occur and what problems they are trying to solve, the better we can design around these problems to avoid them cropping up in the future.

@@

\section{Existing Literature} \label{existing_literature}
This section will introduce the topics of API design and software evolution and provide a brief review of studies that are directly related to our study.

@@

\subsection{Software Evolution} \label{software_evolution}
%%Categories of maintenance and types of maintenance
Software evolution is a field which studies the application of software maintenance activities and processes on software and the resulting, evolved versions of the software. The concept of software maintenance has existed since the 1960s. Well-known studies from this era established the importance of software maintenance \cite{lientz1980software}. One study would go on to have a particularly profound effect on software maintenance by establishing a set of four categories which describe the intents behind change to software \cite{lientz1980software}. These categories were refined in a further study, in which the number of categories were increased from four to twelve \cite{chapin2001types}. The categories proposed were: Enhancive, corrective, reductive, adaptive, performance, preventive, groomative, updative, reformative, evaluative, consultive and training \cite{chapin2001types}. Each category describes a type of software maintenance activity which relate to one of three particular areas; the code, the software and the customer-experienced functionality \cite{chapin2001types}. This study also helped to further the distinction between software maintenance and software evolution. The study suggested that software maintenance is a part of software evolution and that the two are not interchangeable \cite{chapin2001types}.

@@

\section{Case Company Description} \label{case_company_description}
The case company is a medium-sized company in the domain of security cameras. The company has its headquarters in Lund, Sweden, but it operates worldwide. The company is the global market leader in the markets of network cameras and video encoders. The company develops embedded software for the security cameras and some of their products are designed to be accessible through APIs that are also developed by the company.

@@

The case was selected using convenience sampling \cite{flyvbjerg2006five} and a maximum variation strategy was used \cite{benbasat1987case}. As such, the goal was to include as many companies as possible in this study.

@@

\section{Methodology} \label{methodology}
%Intro and summary
This study has been conducted using the case study methodology \cite{runeson2009guidelines}. The goal of the study was twofold. Firstly, the aim was to empirically \textit{explore} the motives behind changes to APIs and secondly, our ambition was to \textit{validate} the results' level of correctness by comparing our findings with the case company's explanation of their motives. Our research can be classified as an embedded case study \cite{yin2013case}, since both the API code and API documentation have been used as units of analysis.

%Reasons for choosing case study
The reason the selected methodology was used is that it is essential to study the phenomenon of API change in its natural context. The applicability of case studies in such scenarios is supported by existing literature \cite{benbasat1987case} \cite{runeson2009guidelines} \cite{yin2013case} \cite{robson2002real}. An alternative approach that was considered was design research, but if a prototype API was to be used instead of one which is tried and tested in an industry setting, the study would lack real-life context \cite{runeson2009guidelines}. An additional motivation for why the case study approach was used is that there is little existing research conducted in the area of motives behind API change and that input from the industry is vital to the success of the research; especially to be able to answer \textbf{RQ2}.

@@

%Description of code analysis
More specifically, the approach undertaken was divided into three steps where the data was first coded based on which category of maintenance \cite{chapin2001types} the motive of the change corresponded to. The data were then mapped and coded to the laws presented by \cite{lehman1980programs}. The codes were then analysed from a high-level point of view, the main goal of which was to identify trends in the codes. The final step involved analysis of these trends and it was during this stage the hypotheses were generated and formulated. The same approach taken to reduce risks of research bias and human error in the data collection was performed during results analysis.

@@

\subsection{Continuing Change}
The case company described that including new functional content was one of the top priorities for the 2.0 release. This is supported by the characteristic of the changes identified in this study. 4 of the 13 changes were additions of new modules and in the re-implemented modules a lot of functional content was added, too. It was also mentioned that the API users had previously requested more frequent updates to the API, despite the drawbacks this could have. One of the main reasons for continuously adapting the API is also to force (!!!wording) the API users to use the API in the way that it is intended. Previously, many applications which worked around the lacking (!!!negative wording) functionality had been developed. These applications were fully functional, but were developed very differently from user to user (!!!wording). We therefore conclude that the law of continuing change holds for APIs.

@@

\subsection{Self Regulation} \label{law3}
When version 1.4 was introduced, no new functionality or re-structuring was introduced. Version 1.4 only added support for additional build platforms. When comparing the changes introduced in 2.0 to the ones previously introduced, it is clear that the growth of the API has increased gradually. During the interview it was also mentioned that the future plans for the API did not include any major changes and that only a few minor changes were planned in the near future. The future updates to the API will also be made incrementally and there will not be a big bang update similar to the one of 2.0. This supports that the law of self regulation applies to APIs, assuming that the current stage of the life-cycle of the API can be considered as the centre of the normal distribution curve of the APIs growth.

\subsection{Conservation of Organisational Stability}
The changes to the ACAP API has been made very sporadically, where only two new versions have been released since the deployment of version 1.0. As mentioned in Section \ref{law3} the changes introduced in version 1.4 did not involve any changes to the existing API functions. Consequently, the only significant update to the API was made in the transition to version 2.0. The internal development of the API has also been made sporadically and there have been periods where the API has not been actively developed. This suggests that the law of conservation of organisational stability does not apply to APIs.

@@

\subsection{Continuing Growth}
Large parts of the changes between the versions were strictly additions of new functional content. 4 new modules were added and significant additions of new functionality were added to the modules which were re-structured from version 1.4. The interviewee also made it clear that increasing the functionality was one of the major goals with the 2.0 update. This also ties in with the ambition of controlling the way the API users use the API, as mentioned previously. It was also mentioned that the case company had received feedback from the API users that more frequent updates to the functional content was desirable, even though this may require them to update their application code. Because of this, we conclude that the law of continuing growth applies to APIs.

@@

\subsection{Feedback System}
One of the future goals of improvement of the case company is to better include the API users and end-users of the applications in their feedback loop. Currently, the only the API users are included in this feedback loop and the API architect expressed the desire to also include the end-users in this loop. The reasoning for this was to increase the quality and to be able to deliver content which is useful for the users and which leads to good applications being developed. One example of how the API users have been included in the feedback loop, was prior to version 2.0 being released, when API users implemented functionality related to sound, even though there was no interface for this in the API. The experiences of this API user, were later considered when developing the axsound module. One benefit of having a more structured feedback loop, would be that the API developers would be more and faster aware of potential errors in the implementation of the API. Another ambition which was mentioned, was to bring scientific theories and best practices from academia into consideration when developing the API. This would aim to complement the current strategy which is largely based around the expertise and experiences of the API architects. Because of these factors and clearly expressed ambitions, we conclude that the law of feedback system is true for APIs.

@@

RQ1.1 - To what extent can we reverse engineer API change decisions?

After cross examining our results with the interviewee and their responses, we found that roughly 78\% of our categorisation matched the actual categorisation of the changes. While it is obvious that we did not manage to successfully categorise every single change, we did categorise a high enough percentage to form a solid conclusion; it is possible to reverse engineer the majority of change decisions using this categorisation technique.

For future studies, this technique could be improved by performing more actions along side the code review. For example, having access to and analysing commit logs, design meetings and other such documentation would allow a much more thorough determination of intents to be collected and analysed.

RQ2
What this means for industry:
	Greater emphasis should be placed upon communicating with the end users / application developers. APIs should be made with maintainability and extendibility in mind as they seem to be what changes the most between versions. They should probably improve their feedback system to get more feedback from the API users.

@@

What this means for academia:
	Academia could benefit from the results as a means for further research - We have found that it's possible to reverse engineer these design decisions from a simple study, so perhaps more advanced studies should be conduct at a larger scale? -- Why though?
